---
title: Wallace Session `r Sys.Date()`
output: html_document
---

```{r, include=FALSE}
library(knitr)
knit_engines$set(asis = function(options) {
  if (options$echo && options$eval) knit_child(text = options$code)
})
knitr::opts_chunk$set(message = FALSE, warning = FALSE, eval = FALSE)
```

```{r, include = FALSE, eval = TRUE}

```

Please find below the R code history from your *Wallace* v1.9.9* session. 

You can reproduce your session results by running this R Markdown file in RStudio. 

Each code block is called a "chunk", and you can run them either one-by-one or all at once by choosing an option in the "Run" menu at the top-right corner of the "Source" pane in RStudio. 

For more detailed information see <http://rmarkdown.rstudio.com>).

### Package installation

Wallace uses the following R packages that must be installed and loaded before starting.
```{r}
library(spocc)
library(spThin)
library(dismo)
library(rgeos)
library(ENMeval)
```

Wallace also includes several functions developed to help integrate different packages and some additional functionality. For this reason, it is necessary to load the file `functions.R`, The function `system.file()` finds this script, and `source()` loads it.

```{r}
source(system.file('shiny/funcs', 'functions.R', package = 'wallace'))
```

Your analyses are below.

## Record of analysis for *`r "{{spName}}"`*.

```{asis}
### Obtain Occurrence Data
```

```{asis, echo = {{occsDb_knit}}, eval = {{occsDb_knit}}, include = {{occsDb_knit}}}
You searched the `r "{{occDb}}"` database for *`r "{{spName}}"`*, limited to `r {{occNum}}` records. 
```

```{r, echo = {{occsDb_knit}}, include = {{occsDb_knit}}}
# query selected database for occurrence records
occs.query <- c1_queryDb(spName = "{{spName}}", occDb = "{{occDb}}", occNum = {{occNum}})
occs <- occs.query$cleaned
```

```{asis, echo = {{occsUser_knit}}, eval = {{occsUser_knit}}, include = {{occsUser_knit}}}
User CSV path with occurrence data. If the CSV file is not in the current workspace, change to the correct file path (e.g. "/Users/darwin/Documents/occs.user").
```

```{r, echo = {{occsUser_knit}}}
# NOTE: provide the path to the .csv file
occs.path <- ""
# get a list of species occurrence data
occs.user <- c1_userOccs(csvPath = occs.path, csvName = "{{userOccsCsvName}}")
occs <- occs[["{{spName}}"]]$cleaned
```

```{asis}
### Process Occurrence Data
```

```{asis, echo = {{removeByID_knit}}, eval = {{removeByID_knit}}, include = {{removeByID_knit}}}
Remove the occurrence localities by ID.
```

```{r, echo = {{removeByID_knit}}, include = {{removeByID_knit}}}
# remove the rows that match the occIDs selected
occs <- c2_removeByID(occs, {{removeByID_id}})
```

```{asis, echo = {{selectByID_knit}}, eval = {{selectByID_knit}}, include = {{selectByID_knit}}}
The following code recreates the polygon used to select occurrences to keep in the analysis.
```

```{r, echo = {{selectByID_knit}}, include = {{selectByID_knit}}}
occs <- c2_selectOccs(occs, "{{selectByID_xy}}", "{{selectByID_id}}")
```
