---
title: "Wallace Vignette: Authoring Modules"
author: "Jamie M. Kass"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{wallace Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  ---
  
  ```{r setup, echo = FALSE, include=FALSE}
library(knitr)
knitr::opts_chunk$set(collapse=TRUE, message=FALSE, eval = FALSE, warning=FALSE, comment="#>")
```

- [Introduction](#intro)
- [Structure of `wallace`](#structure)



# Introduction {#intro}

[*Wallace*](https://wallaceecomod.github.io/) is a `shiny` application that allows users to build and evaluate models of species niches and distributions from scratch using a point-and-click interface (link to [CRAN](https://cran.r-project.org/package=wallace))). The application features an interactive map and dynamic plots and tables. Data for the models can be download from online databases or uploaded by the user, and most results can be downloaded, including the option to save R code that can reproduce your analysis. 

This vignette walks the user through the process of building a new module (explained below). Researchers who want to share their techniques with the user community can build a module and submit it via a pull request to the [*Wallace* Github page](https://github.com/wallaceEcoMod/wallace). 

### Components {#components}

In *Wallace*, "components" are steps of a species distribution/niche modeling analysis. These components were decided upon by the development team, but by no means are exhaustive -- we envision more may be added in the future. There are currently 8 components that proceed sequentially:

1: Obtain Occurrence Data **->** 2: Process Occurrence Data **->** 3: Obtain Environmental Data **->** 4: Process Environmental Data **->** 5: Partition Occurrence Data **->** 6: Model **->** 7: Visualize Model Results **->** 8: Project Model

### Modules {#modules}

Within components, there are "modules", which represent options for each analysis step. For example, within component 6 "Model", the modules represent different distribution/niche model choices. *Wallace* comes with built-in modules, but users can build their own modules to extend the app's functionality. Some users may have a different preferred way of approaching some part of the analysis, and by authoring a module they can share this method with the user community.


## Structure of `wallace`

As mentioned above, `wallace` was written using the web application development package `shiny`. It is highly recommended you go through the `shiny` [tutorial](http://shiny.rstudio.com/tutorial/) and the `shiny` module [info page](https://shiny.rstudio.com/articles/modules.html) before proceeding in order to understand the basics.

All the relevant code is in **inst/shiny**. The user interface and server code are in separate scripts, **ui.R** and **server.R**. The behavior of the user interface is governed by **ui.R**, and the analysis code is run by **server.R**. Additional functions used throughout are stored in **/funcs/functions.R**. Modules are all separate scripts, and are stored in **/modules**. Modules have the following naming convention: *mod_cNUMBER_NICKNAME*, where the module called "Query Database" for component 1 is named *mod_c1_queryDB*.

Each component has a set of radio buttons that allow users to choose between modules. What follows are conditional panels that will display each module if it is selected. Within **ui.R**, you can find each component organized this way:

```{r}
conditionalPanel("input.tabs == 1",
                 h4("Obtain Occurrence Data"),
                 radioButtons("occSel", "Modules Available:",
                              choices = list("Query Database" = 'db', "User-specified Occurrences" = 'user'),
                              selected = 'db'),
                 HTML('<hr>'),
                 conditionalPanel("input.occSel == 'db'",
                                  div(paste('Module: Query Database'), id="mod"),
                                  uiTop('spocc', 'Interface to Species Occurrence Data Sources'),
                                  HTML('<hr>'),
                                  queryDB_UI('c1_queryDB'),
                                  actionButton("goDbOccs", "Query Database"),
                                  strong("Download database occurrence localities (.csv)"), br(), br(),
                                  downloadButton('dlDbOccs', "Download DB Occurrences"),
                                  HTML('<hr>'),
                                  uiBottom('spocc', "Scott Chamberlain, Karthik Ram, Ted Hart")
                 ),
                 conditionalPanel("input.occSel == 'user'",
                                  div('Module: User-specified Occurrences', id="mod"),
                                  HTML('<hr>'),
                                  userOccs_UI('c1_userOccs'),
                                  actionButton("goUserOccs", "Load Occurrences")
                 )
),
```

